draft 1

### 2016-07-26 finalwriteup/draft.rmd
finished comparison table

@@

ANOVA is closely related to linear fit models but incorporates the analysis of differences in group means (i.e. categorical variables) [@Lunney1970]. Since the method assumes variables are normally distributed, numeric variables must be normalised before analysis to achieve a statistical distribution as close to Gaussian as possible. Therefore a disadvantage of this method is that linear relationships and normally distributed variables do not necessarily represent real world data accurately [@Breiman2001b].

In order to compare variable importance, two ANOVA outputs were assessed: the linear coefficients for each variable and the p-values for the coefficients [@Markham2016]. The coefficient does not indicate importance relative to the other variables because its value reflects values within the variable. The sign of the coefficient however indicates whether the variable is directly or inversely related to the response variable. A significance test for whether the coefficient equals zero is performed on each coefficient and the p-value is reported. Therefore, if the null hypothesis is rejected (p value < 0.05), changes in the explanatory variable are indeed related to changes in the response variable. P-values are a good indication of whether variables have a relationship with the target variable, however they do not necessarily rank importance. Precision of variable values can influence p-values, for example a variable with high precision will have a smaller p-value, whereas a variable measured roughly will have a higher p-value. This does not necessarily make the more precisely measured variable more important than the other. It is also important to keep in mind that, with a dozen variables, chance alone can produce a variable with a p-value < 0.001 7% of the time [@Rice1989].

@@

The Random Forest algorithm can produce a permutation variable importance for each variable as part of its output. Variable importance is represented as a score, which is calculated by performing permutations on each variable in the test data for each tree, then re-running the test data through the trees. More specifically, consider a single variable. The values of this variable in the out-of-bag sample (the test data that the Random Forest algorithm partitions before building each tree) are randomly permuted (reordered) for each tree. The purpose of permuting is to mimic the absence of that variable. The trees in the Random Forest are then run again. Variable importance can now be derived by comparing the results of the forest where the covariate has been permuted against the original Random Forest. If the accuracy of the permuted forest is much less than the original forest, then that variable was important. Permutation importance scores are calculated as the mean decrease in accuracy over all trees for the permuted variable in the model [@Breiman2005].

These scores provide valuable insight because they represent the impact of each predictor variable individually as well as in multivariate interactions with other predictor variables [@Strobl2007]. However, a severe disadvantage is that the scores are not reliable when the variables differ in their numeric scales or their counts of categories. Also, the importance of correlated predictors is overestimated and the algorithm tends to favour variables that have many possible splits or many missing values [@Strobl2007]. The intuition behind this is that if a variable has more points to split the data, there are more opportunities for it to split the response variable favourably. However this does not necessarily indicate the variable impacts the response variable the most. Overall, the permutation importance is useful to compare variables in a non-linear environment and can include multivariate interactions, but overestimates the importance of variables with with many categories and some numeric variables.

@@

The procedure behind the significance test at each split in a cTree is as follows. To determine which explanatory variable should be used at each split, each variable is permuted in every possible way, and a correlation value is calculated between the tested variable and response variable, for each permutation. The unchanged variable correlation is then compared with the correlation values for all permutations of that variable. From this, a p-value for the true correlation value compared to the permuted correlation values can be calculated. The predictor variable with the lowest p-value is then selected as the splitting variable [@Hothorn2006]. Using an ensemble forest of these trees, variable importance is then calculated in the same fashion as Random Forests, via permuting a variable, re-running the forest and comparing the decrease in response variable accuracy. Several sources recommend cForests variable importance methods because its statistical p-value tests at each split removes the bias present in other tree ensemble variable importance outcomes such as Random Forests (as discussed) and Boosted Trees [@Strobl2007][@Strobl2009][@Hothorn2006].

@@

@@

Several research groups involved in the high profile data science competition, Netflix Prize, developed sophisticated methods of model stacking, otherwise known as ensemble methods or model blending [@Sill2009]. In this document, the term 'blending' will be used to avoid confusion with ensemble tree methods. The idea behind model blending is to combine predictions from several models derived from different methods that have different theoretical foundations. In this way, the strengths of each method can be combined. Historically, in statistics this was called model averaging and since the 1990's research has shown that averaging results from different methods provides better predictive accuracy than any single model [@Madigan1994]. New methods of model blending were developed for the Netflix Prize that applied sophisticated machine learning techniques to big socially generated data. The prize-winning solution was a complex blend of sub-blends which interacted results of individual models with the original variables. When original variables are used in blending they are called meta-features [@Sill2009]. The multiple layers of blending gave incremental improvements in predictive accuracy, but the computational cost of all the complex layers did not justify the benefits for Netflix in practice and a simplified model was adopted. Therefore, in this project just a single layer of blending was trialed.

>>>>

### Summary Comparison between the Construction and IT Industry

As previously stated, there are many parallels and some contrasts in cost estimation research in the Construction industry vs. the IT industry. These have been summarised into a table below, which is coloured according to similarities and differences.

![Summary table of similarities and differences between cost estimation research in the IT and Construction industry](images/lit_review/compare_tbl.png)

@@

ANOVA is closely related to linear fit models but incorporates the analysis of differences in group means (i.e. categorical variables) [@Lunney1970]. It is good practice to normalise numeric variables before analysis to minimise effects of any high leverage data points. ANOVA assumes linear relationships between explanatory variables and the response variable, which can be a disadvantage as real world data is not necessarily linear [@Breiman2001b].

In order to compare variable importance, two ANOVA outputs were assessed: the linear coefficients for each variable and the p-values for the coefficients [@Markham2016]. The coefficient does not indicate importance relative to the other variables because its magnitude reflects values within the variable. The sign of the coefficient however indicates whether the variable is directly or inversely related to the response variable. A significance test for whether the coefficient equals zero is performed on each coefficient and the p-value is reported. Therefore, if the null hypothesis is rejected (p value < 0.05), changes in the explanatory variable are indeed related to changes in the response variable. P-values are a good indication of whether variables have a relationship with the target variable, however they do not necessarily rank importance. Precision of variable values can influence p-values; for example a variable with high precision will have a smaller p-value, whereas a variable measured roughly will have a higher p-value. This does not necessarily make the more precisely measured variable more important than the other. It is also true that with a dozen variables, chance alone can produce a variable with a p-value < 0.001 7% of the time [@Rice1989].

@@

The Random Forest algorithm can produce a permutation variable importance for each variable as part of its output. Variable importance is represented as a score, which is calculated sequentially for each variable by comparing predictive accuracy between out-of-bag (OOB) data run through the trees without, then with a variable permutation (reordering). Since each tree is a bootstrapped sample of the complete dataset, approximately 1/3 of the data is held aside as the OOB sample for each tree. It is passed through its partner tree and produces an unbiased estimate of error and in this case, is also used to estimate variable importance. After the original OOB sample has been run through the trees, a variable is permuted in all OOB samples and are run through all of the trees again.

The purpose of permuting is to mimic the absence of that variable. Now the accuracy of the trees where the covariate has been permuted can be compared against the original Random Forest. For a categorical response variable, refer to the equation describing variable importance score below:

$Raw\ VI(X_i) = \frac{\sum_{n = 1}^{ntrees} \left(C_{n} - C_{n,X_1} \right)}{ntrees}$

Where

$$
\begin{aligned}
Raw VI(X_i) &= \textrm{raw variable importance for variable}\ X_i \\
ntrees &= \textrm{number of trees specified in the Random Forest} \\
C_n &= \textrm{Number of correctly classified cases in the nth tree using the OOB sample} \\
C_{n,X_1} &= \textrm{Number of correctly classified cases in the nth tree using the OOB sample with permuted} \\
&\qquad \textrm{values for} X_i
\end{aligned}
$$

If the accuracy of the permuted forest is much less than the original forest, then that variable was important. As shown in the equation, permutation importance scores are calculated from the mean decrease in accuracy over all trees for the permuted variable in the OOB samples [@Breiman2005].

These scores provide valuable insight because they represent the impact of each predictor variable individually as well as in multivariate interactions with other predictor variables [@Strobl2007]. However, a severe disadvantage is that the scores are not reliable when the variables differ in their numeric scales or their counts of categories. Also, the importance of correlated predictors is overestimated and the algorithm tends to favour variables that have many possible splits or many missing values [@Strobl2007]. The intuition behind this is that if a variable has more points to split the data, there are more opportunities for it to split the response variable favourably. However this does not necessarily indicate the variable predicts the response variable the most. Overall, the permutation importance is useful to compare variables in a non-linear environment and can include multivariate interactions, but overestimates the importance of variables with many categories and some numeric variables.

@@

The procedure behind the significance test at each split in a cTree is as follows. To determine which explanatory variable should be used at each split, each variable is permuted in every possible way, and a correlation value is calculated between the tested variable and response variable, for each permutation. The unchanged variable correlation is then compared with the correlation values for all permutations of that variable. From this, a p-value for the true correlation value compared to the permuted correlation values can be calculated. The predictor variable with the lowest p-value is then selected as the splitting variable [@Hothorn2006]. Using an ensemble forest of these trees, variable importance is then calculated in the same fashion as Random Forests, via permuting a variable in the OOB sample, re-running the forest and comparing the decrease in response variable accuracy. Several sources recommend cForests variable importance methods because its statistical p-value tests at each split removes the bias present in other tree ensemble variable importance outcomes such as Random Forests (as discussed) and Boosted Trees [@Strobl2007][@Strobl2009][@Hothorn2006].

@@

The RMSE is a popular method for gauging numeric predictions and is represented by the equation below:

$RMSE =  \sqrt{\frac{\sum_{t = 1}^{n} (\widehat{y}_t - y_t)^2}{n}}$

Where

$$
\begin{aligned}
n &= \textrm{number of data points} \\
\widehat{y} &= \textrm{predicted values of the response variable} \\
y &= \textrm{observed values of the response variable} \\
\end{aligned}
$$

@@

Several research groups involved in the high profile data science competition, Netflix Prize, developed sophisticated methods of model stacking, otherwise known as ensemble methods or model blending [@Sill2009]. In this document, the term 'blending' will be used to avoid confusion with ensemble tree methods. The idea behind model blending is to combine predictions from several models derived from different methods that have different theoretical foundations. In this way, the strengths of each method can be combined. Historically, in statistics this was called model averaging and since the 1990's research has shown that averaging results from different methods provides better predictive accuracy than any single model [@Madigan1994].

New methods of model blending were developed for the Netflix Prize that applied sophisticated machine learning techniques to big socially generated data. The prize-winning solution was a complex blend of sub-blends which interacted results of individual models with the original variables. When original variables are used in blending they are called meta-features [@Sill2009]. The multiple layers of blending gave incremental improvements in predictive accuracy, but the computational cost of all the complex layers did not justify the benefits for Netflix in practice and a simplified model was adopted. Therefore, in this project just a single layer of blending was trialed.
